Communication in Agile Inter-team Development

### 2019-01-21 Draft/Parts/Conclusion.tex
Many typo fixes, tone corrects, and clarifications.

\section{Conclusion}\label{sec:conclusion}
As we have seen, inter-team communication can be a significant obstacle when agile teams work together, as intermediaries in communication and lack of personal communication causes a disruption in the agile process.
To reduce this problem a number of small adaptions can be made to the process to ensure the flow of information.
However, the most important adaption is a well defined design and architecture which is available across teams and enables teams to produce individual system components which can be easily integrated.

>>>>

\section{Conclusion}\label{sec:conclusion}
As we have seen, inter-team communication can be an obstacle when agile teams work together, as intermediaries in communication and lack of personal communication causes a disruption in the agile process.
To reduce this problem a number of adaptions can be made to the process to ensure the flow of information.
However, the most important adaption is a well defined design and architecture which is available across teams and enables teams to produce individual system components which can be easily integrated.

### 2019-01-21 Draft/Parts/Introduction.tex
Minor adjustment in intro.
Many typo fixes, tone corrects, and clarifications.

%AM: What is the problem?
%introduce the need to scale to multiple agile teams
Agile approaches are popular in software development for numerous reasons.
However, the structure of the agile process does not lend itself to large teams and is not suitable for large projects.
With the growth in complex software, situations arise where multiple agile teams need to work together to achieve a common goal.
%thesis statement
When scaling to parallel agile teams, the most important consideration is quality communication, both personal and architectural.
%
For an example of a company developing a complex piece of software, consider Spotify AB who has over 30 teams \cite{kniberg12}.
Spotify is a streaming music client which supports media, commercial, and social features.
The basic element of development organization at Spotify is the Squad, analogous to a Scrum team.
%Each Squad is the autonomous developer, with a localized working area, of a particular sub-system in the product.
Squads in related areas are grouped together into Tribes. 
The overall structure is illustrated in Figure~\ref{fig:spotify_structure}.
\begin{figure}[h]
  \includegraphics[width=\linewidth]{images/kniberg12_structure.png}
  \caption{Agile team organization at Spotify \cite{kniberg12}.}
  \label{fig:spotify_structure}
\end{figure}

%AM: Why is it interesting/important?
%introduce agile
As outlined in the Agile Manifesto \cite{beck2001agile}, an agile process gives individuals and interactions value over processes and tools.
An agile team should contain all that is required for them to do their work, thus interactions are intensely intra-team.
However, diverse feature sets such as required in the Spotify application necessitate many skillsets and substantial manpower.
By moving to a multi-team model, complex projects can benefit from agility, provided that the process is adapted.
%personal communication
Multiple teams increases the importance of communication, from simple interpersonal communication to system specification. 
The agile approach of interacting in-person can fail due to reasons such as the vertical structure of the company organization and layers of middle managers \cite{dzone_article}.
The large number of intermediaries can cause an agile process to breakdown as it takes longer than usual time for messages to reach their destination than can afforded.
%arch communication
In an agile process, where Big Design Up-front may typically be avoided, the move to multiple teams motivates a need to communicate design decisions.
%todo: component boundaries?
For successful integration of technological components, system elements must have well defined functional specifications, and interfaces.

@@

%outline of rest of paper
This document is organized as follows: Introduction, Communication Concerns, Supporting Communication in Inter-team Development, and Conclusion. 
In Section ~\ref{sec:spt_ex}, we discuss example communication concerns in an agile project.
In Section ~\ref{sec:prop_appro}, we present some process adaptations that can help to resolve issues.
Lastly, Section ~\ref{sec:conclusion} concludes with an overview of important aspects of inter-team agile development.

>>>>

%AM: What is the problem?
%introduce the need to scale to multiple agile teams
Agile approaches are popular in software development for many reasons.
However, the structure of an agile process does not lend itself to large teams and is not suitable for large projects.
In complex software, situations arise where multiple agile teams need to work together to achieve a common goal.
%thesis statement
When scaling to parallel teams, the most important consideration is effective communication, both personal and architectural.
%
For an example of a company developing a complex piece of software, consider Spotify AB who has over 30 teams \cite{kniberg12}.
Spotify is a streaming music client which supports media, commercial, and social features.
The basic element of development organization at Spotify is the Squad, analogous to a Scrum team.
%Each Squad is the autonomous developer, with a localized working area, of a particular sub-system in the product.
Squads in related areas are grouped together into Tribes. 
The overall structure is illustrated in Figure~\ref{fig:spotify_structure}.
\begin{figure}[h]
  \includegraphics[width=\linewidth]{images/kniberg12_structure.png}
  \caption{Agile team organization at Spotify \cite{kniberg12}.}
  \label{fig:spotify_structure}
\end{figure}

%AM: Why is it interesting/important?
%introduce agile
As outlined in the Agile Manifesto \cite{beck2001agile}, an agile process gives individuals and interactions value over processes and tools.
An agile team should contain all that is required for them to do their work, thus interactions are intensely intra-team.
However, diverse feature sets such as required in the Spotify application necessitate many skill sets and substantial manpower.
By moving to a multi-team model, large projects can benefit from agility, provided that the process is adapted.
%personal communication
Multiple teams increases the importance of communication, from simple interpersonal communication to system specification. 
The agile approach of interacting in-person can fail due to reasons such as the vertical structure of company organization and layers of middle managers \cite{dzone_article}.
The large number of intermediaries can cause an agile process to breakdown as it takes longer than usual for messages to reach their destination than can afforded.
%arch communication
In an agile process, where Big Design Up Front may typically be avoided, the move to multiple teams motivates a need to communicate design decisions.
%todo: component boundaries?
For successful integration of technological components, system elements must have well defined functional specifications, and interfaces.

@@

%outline of rest of paper
This document is organized as follows: Introduction, Communication Concerns, Supporting Communication in Inter-team Development, and Conclusion. 
In Section ~\ref{sec:spt_ex}, we discuss example communication concerns.
In Section ~\ref{sec:prop_appro}, we present some process adaptations that can help to resolve issues.
Lastly, Section ~\ref{sec:conclusion} concludes with an overview of important aspects of inter-team agile development.

### 2019-01-21 Draft/Parts/Main.tex
Many typo fixes, tone corrects, and clarifications.

\section{Communication Concerns}\label{sec:spt_ex} 
%overview
A fundamental issue with tasking teams with different goals is that teams can become isolated with respect to the context of their work.
% problem
Consider a scenario at Spotify \cite{kniberg12}: a tester on a particular agile team invested time to solve a particular problem, which was also faced by another tester on a different team. Without communication, these team members would perform redundant work.
To address this problem Spotify defines Chapters, which are crosscutting divisions which gathers individuals with similar roles across different Squads. 
At Spotify \cite{kniberg12}, offices are laid out so that the Squads making up a Tribe are spatially close. This leads to an environment where an informal exchange of information on the Tribe's work will occur.
%Although larger than Squads (7-10 members), Tribes are limited in size (< 100) to ensure smooth communication.
%todo: any more issues? Dunbar

In general, consistency becomes as issue when expanding to multiple teams.
The product owner in an agile process model is the person with the project vision and if it is a small team the product owner can clearly communicate this to every team member.
Members can even query the product owner directly when in doubt.
However, in large scale projects where multiple teams are working together, it is practically impossible for the owner to keep doing what they did in a small team.
That’s where a formal definition of their vision is useful as it gives teams something to consult to when encountering a design decision.

\section{Supporting Communication in Inter-team Development}
\label{sec:prop_appro} 
The general problem of communication can be tackled from two perspectives.
First, stakeholders at and across different levels need to have communication channels.
This can be addressed by adaptations in development process and leveraging technology.
Second, the quality (e.g., accuracy) of communication needs to be high.
This can be addressed by thorough system specification through architecture and design.

@@

%Changes in standups
One critical aspect of an agile process is daily standups which are a platform for letting each team member know the work being done in other parts of the team and correlate it with the work they are performing.
This uncovers differences in understanding early in development and prevents last minute discovery of mismatches in development.
In the case of multiple agile teams this problem is compounded as a stand up is a closed activity of the team itself, thus preventing other teams from knowing the results or discussions of each other.
This can be resolved by having a representative of each concerned team be present in the daily standup thus letting each team know the status of the other teams.

%Changes to Product Owner
Although the product owner is responsible for agile teams under his supervision, projects with multiple agile teams may have multiple product owners, and over time the vision of the owners may drift and lead the development in different directions.
This can be limited by having regular meetings of product owners where the scope and vision of the project can be synchronized.
%This can be a bi-weekly or monthly meeting depending on the size of the project.

%Changes in Planning Sessions
All the initial, intermediate and final planning sessions should be made such that teams which are or could be impacted by that part of the project are part of the meeting.
This can assist in early agreement on high-level requirements and standardization of inter-team interfaces.

@@

%Communication Tools
Since agile focuses on personal interaction, particularly face-to-face conversation, video conferencing tools can lower the bar to inter-team interactions.
They also make communication real-time, reducing delay due to time spent in communicating ideas across teams.
%Integration Tools
Continuous integration tools can go a long way in identifying inter-team problems early in development
as it reveals the impact of changes on the entire project.

@@

System design or architecture acts as a common vision for not only the teams but also for the communication between all the product owners.
The system design acts as a “deadlock breaker in decisions” \cite{architecureRole_article} for situations where teams disagree on a particular design path.
%non specification fix
A second approach addresses this issue from an organizational standpoint by defining a "system owner" role \cite{kniberg12}.
This role is more casual than an architect role, and focuses on defining a "go-to" person who can maintain long term stewardship over a sub-system.

>>>>

\section{Communication Concerns}\label{sec:spt_ex} 
%overview
A fundamental issue with tasking teams with different goals is that teams can become isolated with respect to the context of their work.
% problem
Consider a scenario at Spotify \cite{kniberg12}: a tester on a particular agile team invested time to solve a particular problem, which was also faced by another tester on a different team. Without communication, these team members would perform redundant work.
To address this problem Spotify defines Chapters, which are crosscutting groups which gather individuals with similar roles across different Squads. 
At Spotify \cite{kniberg12}, offices are laid out so that the Squads making up a Tribe are spatially close. This leads to an environment where an informal exchange of information on the Tribe's work will occur.
%Although larger than Squads (7-10 members), Tribes are limited in size (< 100) to ensure smooth communication.
%todo: any more issues? Dunbar

In general, consistency becomes an issue when expanding to multiple teams.
The product owner in an agile process model is the person with the project vision, and with a small team, the product owner can clearly communicate this to every team member.
Members can even query the product owner directly when in doubt.
However, in large projects where multiple teams are working together, it is practically impossible for the owner to continue what they did in a small team.
This motivates a need for a formal definition of their vision as it gives teams something to consult to when encountering a design decision.

\section{Supporting Communication in Inter-team Development}
\label{sec:prop_appro} 
The general problem of communication can be tackled from two perspectives.
First, stakeholders at and across different levels need to have communication channels.
This can be addressed by adaptations in development process and leveraging technology.
Second, the quality (e.g., accuracy) of communication needs to be high.
This can be addressed by thorough system specification through architecture.

@@

%Changes in standups
One critical aspect of an agile process is daily standups which are a platform for letting each team member know the work being done in other parts of the team and correlate it with the work they are performing.
This uncovers differences in understanding early in development and prevents last minute discovery of discrepancies.
In the case of multiple agile teams this problem is compounded as a stand up is a closed activity of the team itself, thus preventing other teams from knowing the results or discussions of each other.
This can be resolved by having a representative of each concerned team be present in the daily standup thus letting each team know the status of the other teams.

%Changes to Product Owner
Although the product owner is responsible for agile teams under his supervision, projects with multiple agile teams may have multiple product owners, and so over time the vision of the owners may drift and lead the development in different directions.
This can be limited by having regular meetings of product owners where the scope and vision of the project can be synchronized.
%This can be a bi-weekly or monthly meeting depending on the size of the project.

%Changes in Planning Sessions
All the initial, intermediate and final planning sessions should be organized such that teams which are, or could be, impacted by that part of the project are part of the meeting.
This can assist in early agreement on high-level requirements and standardization of inter-team interfaces.

@@

%Communication Tools
Since agile focuses on personal interaction, particularly face-to-face conversation, video conferencing tools can lower the bar to inter-team interactions.
They also make communication real-time, reducing delay due to time spent communicating ideas across teams.
%Integration Tools
Continuous integration tools can go a long way in identifying inter-team problems early in development
as it reveals the impact of changes on the entire project.

@@

System design and architecture acts as a common vision for not only teams but also for the communication between all the product owners.
The system design acts as a “deadlock breaker in decisions” \cite{architecureRole_article} for situations where teams disagree on a particular design path.
%non specification fix
A second approach addresses this issue from an organizational standpoint by defining a "system owner" role \cite{kniberg12}.
This role is more casual than an architect role, and focuses on defining a "go-to" person who can maintain a long term stewardship over a sub-system.
